---
title: "Status Report - Comparison of PAMS and Banks Equity Ownership"
author: "Jacob Kastl"
date: \today
output: 
  html_document:
      toc: true
      toc_depth: 2
      theme: flatly
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Error', '\u2716 **Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Warning:', '\u2716 **Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('\n\n<div class="alert alert-success">',
           gsub('##', '\n \u2714', x),
           '</div>', sep = '\n')
   }
)


config <- config::get()

# set configs
pams_path <- config$compare_pams_banks_eo$paths$path_pams
banks_eo_path <- config$compare_pams_banks_eo$paths$path_banks_eo

library(dplyr)
library(tibble)
library(readxl)
library(tidyr)
library(glue)
library(pacta.data.preparation)
```

## Introduction

This report checks if the PAMS and Banks Equity Ownership (EO) datasets of a common data release are consistent. For that, it compares two datasets: `pams` and `banks_eo` to ensure data consistency.

In general, it is expected that the Banks Equity Ownership dataset can be recreated from the PAMS dataset. Beyond minor differences due to rounding, we would expect that it is possible to derive the Banks EO dataset from PAMS after several steps of data wrangling.

We will check if the above expectation holds and gather any incosnsistencies and unexpected patterns.

## Load Data

The PAMS dataset of the 2023Q2 data release can use the import function `pacta.data.preparation::import_ar_advanced_company_indicators()`, which will return a tidy combination of the sheets contained in the PAMS xlsx. The Banks EO file only requires loading one sheet from an xlsx.

```{r import, echo = FALSE}
comparison_pams <- pacta.data.preparation::import_ar_advanced_company_indicators(pams_path, fix_names = TRUE)
comparison_banks_eo <- readxl::read_xlsx(banks_eo_path, sheet = "Company Indicators - PACTA Comp")
```

## Data Overview

Provide a brief overview of both datasets:

``` {r summary, echo = FALSE}

# Overview of pams
cat("\nSummary of pams:\n")
print(head(comparison_pams))

# Overview of banks_eo
cat("Summary of banks_eo:\n")
print(head(comparison_banks_eo))

```

## Data Preparation
We will now wrangle the PAMS dataset to match the Banks EO dataset as closely as possible.

### Step 1: Correct columns and format
We keep and rename the relevant columns from PAMS needed for Banks EO. And we add a column for the `ald_timestamp`, which is not given in PAMS.

``` {r columns_and_format, echo = FALSE}
ald_timestamp <- unique(comparison_banks_eo$ald_timestamp)
if (length(ald_timestamp) > 1) {
  stop("Cannot derive the appropriate value for `ald_timestamp` from more than one value.")
}

pams_to_banks_eo <- comparison_pams %>%
  dplyr::select(
    dplyr::all_of(
      c(
        "company_id",
        name_company = "company_name",
        "lei",
        is_ultimate_owner = "is_ultimate_parent",
        sector = "asset_sector",
        technology = "asset_technology",
        technology_type = "asset_technology_type",
        plant_location = "asset_country",
        "year",
        "activity_unit",
        "consolidation_method",
        "value_type",
        "value"
      )
    )
  ) %>%
  dplyr::filter(
    .data$consolidation_method == "Equity Ownership"
  ) %>%
  dplyr::select(-"consolidation_method") %>%
  dplyr::mutate(ald_timestamp = .env$ald_timestamp)

```

### Step 2: Formatting to lower case and filter appropriate values

- We format relevant columns to lower case.
- We only keep the years that are given in the Banks EO dataset.
- We only keep power capacity, not power generation.
- We only keep emission intensity values per unit of economic activity, not total emissions.


``` {r format_and_filter, echo = FALSE}
banks_eo_years <- unique(comparison_banks_eo$year)

pams_to_banks_eo <- pams_to_banks_eo %>%
  dplyr::mutate(
    name_company = tolower(.data$name_company),
    sector = tolower(.data$sector),
    technology = tolower(.data$technology),
    technology_type = tolower(.data$technology_type)
  ) %>%
  dplyr::filter(.data$year %in% banks_eo_years) %>% 
  dplyr::filter(!(.data$sector == "power" & .data$activity_unit == "MWh")) %>%
  dplyr::filter(
    .data$value_type == "production" | (.data$value_type == "emission_intensity" & grepl("/", .data$activity_unit))
  )

```

### Step 3: Wrangle production and emission factors into expected format

- `production`, `production_unit`, `emission_factor`, and `emission_factor_unit` are wrangled into wide variables
- `emission_factor` for any sector that is not evaluated using the SDA approach is set to `NA`

``` {r prod_ef_wide, echo = FALSE}
pams_to_banks_eo_production <- pams_to_banks_eo %>%
  dplyr::filter(.data$value_type == "production") %>%
  dplyr::select(-"value_type") %>%
  dplyr::rename(
    production_unit = "activity_unit",
    production = "value"
  )

pams_to_banks_eo_emission_intensity <- pams_to_banks_eo %>%
  dplyr::filter(.data$value_type == "emission_intensity") %>%
  dplyr::select(-"value_type") %>%
  dplyr::rename(
    emission_factor_unit = "activity_unit",
    emission_factor = "value"
  ) %>%
  dplyr::mutate(
    emission_factor = dplyr::if_else(
      .data$sector %in% c("aviation", "cement", "steel"),
      .data$emission_factor,
      NA_real_
    )
  )

pams_to_banks_eo_merged <- pams_to_banks_eo_production %>%
  dplyr::full_join(
    pams_to_banks_eo_emission_intensity,
    by = c("company_id", "name_company", "lei", "is_ultimate_owner", "sector", "technology", "technology_type", "plant_location", "year", "ald_timestamp")
  )

```

### Step 4: Map and Filter Sectors & Technologies

- Some sector and technology names need to be renamed and/or mapped to standard PACTA definitions.
- Shipping sector is removed
- Natural Gas Liquids is removed from Oil & Gas sector
- Freight is removed from Aviation sector

``` {r map_filter_sectors_technologies, echo = FALSE}

pams_to_banks_eo_merged <- pams_to_banks_eo_merged %>%
  dplyr::mutate(
    sector = dplyr::case_when(
      .data$sector == "ldv" ~ "automotive",
      .data$sector == "oil&gas" ~ "oil and gas",
      TRUE ~ .data$sector
    )
  ) %>%
  dplyr::mutate(
    technology = dplyr::case_when(
      .data$sector == "coal" ~ "coal",
      .data$technology %in% c(
        "hybrid no-plug", "ice cng", "ice diesel", "ice e85+", "ice gasoline", "ice hydrogen", "ice propane"
      ) ~ "ice",
      .data$technology == "hybrid plug-in" ~ "hybrid",
      .data$technology == "fuel cell" ~ "fuelcell",
      .data$technology == "oil and condensate" ~ "oil",
      TRUE ~ .data$technology
    )
  ) %>%
  dplyr::filter(
    .data$sector != "shipping",
    !(.data$sector == "oil and gas" & .data$technology == "natural gas liquids"),
    !(.data$sector == "aviation" & .data$technology == "freight")
  )


```

### Step 5: Aggregate

- We aggregate production and emission factors to the company-technology-country-year level.
  - We sum up production
  - We take a weighted mean of the emission factors, using the porduction for weighting


```{r aggregate, echo=FALSE}
pams_to_banks_eo_aggregated <- pams_to_banks_eo_merged %>%
  dplyr::summarise(
    emission_factor = stats::weighted.mean(.data$emission_factor, w = .data$production, na.rm = TRUE),
    production = sum(.data$production, na.rm = TRUE),
    .by = c(
      "company_id",
      "name_company",
      "lei",
      "is_ultimate_owner",
      "sector",
      "technology",
      "plant_location",
      "year",
      "production_unit",
      "emission_factor_unit",
      "ald_timestamp"
    )
  )

```

### Step 6: Accomodate Banks EO particularieties

The Banks EO data set has a few less than obvious traits that need to be accounted for when replicating it from PAMS:

- For sectors with technology pathways, `emission_factor` is set to `NA`.
- We remove combinations of company/technology/country that have zero production values over the entire given timeframe.
- For combinations of company-technology-country that have missing rows in leading years, but positive `production` values in later years, we add rows for the missing leading years and set `production = 0` and `emission_factor = NA_real_`.

```{r particular_wrangling, echo=FALSE}
pams_to_banks_eo_complete <- pams_to_banks_eo_aggregated %>%
  dplyr::mutate(
    emission_factor = dplyr::if_else(
      .data$sector %in% c("automotive", "coal", "hdv", "oil and gas", "power"),
      NA_real_,
      .data$emission_factor
    )
  ) %>%
  dplyr::mutate(
    all_zero = dplyr::if_else(sum(.data$production, na.rm = TRUE) == 0, TRUE, FALSE),
    .by = c("company_id", "lei", "is_ultimate_owner", "sector", "technology", "plant_location", "production_unit", "emission_factor_unit", "ald_timestamp")
  ) %>%
  dplyr::filter(!.data$all_zero) %>%
  dplyr::select(-"all_zero")%>%
  tidyr::complete(
    .data$year,
    tidyr::nesting(
      company_id,
      name_company,
      lei,
      is_ultimate_owner,
      sector,
      technology,
      plant_location,
      production_unit,
      emission_factor_unit,
      ald_timestamp
    ),
    fill = list(
      production = 0,
      emission_factor = NA_real_
    )
  )

```

## Compare

Finally, we compare how closely the replicated data set matches the Banks EO dataset we received.

We compare across the following dimensions:

- How many rows are successfully replicated?
- How many rows are given in the PAMS-based replica, but not in the Banks EO dataset?
- How many rows are given in the Banks EO dataset, but not in the PAMS-based replica?
- What are the patterns of those rows that were not successfully replicated?
- Are the production and emission_factor values inequal between the successfully replicated rows and the Banks EO dataset?
- If not, how far off are they?

Note: In assessing the differences, we ignore the differences in names for now, as the formatting steps are not 100% clear beyond changing the names to lower case. We thus only focus on `company_id` for comparison for now.

```{r successfully_replicated, echo=FALSE}
common_data <- comparison_banks_eo %>%
  dplyr::inner_join(
    pams_to_banks_eo_complete,
    by = c(
      "company_id", "lei", "is_ultimate_owner", "sector", "technology", "plant_location", "year", "production_unit", "emission_factor_unit", "ald_timestamp"
    )
  ) %>%
  dplyr::mutate(
    prod_diff = (production.x - production.y) / production.x,
    ef_diff = (emission_factor.x - emission_factor.y) / emission_factor.x
  )

share_rows_replicated <- nrow(common_data) / nrow(comparison_banks_eo)

```

We calculate the percentage difference in production and emission_factor values between the two datasets. We expect that percent difference to be zero or very close to zero (as there are sometimes rounding errors).

```{r plots_successfully_replicated, echo=FALSE}
common_data %>%
  ggplot2::ggplot(ggplot2::aes(x = prod_diff, y = production.x)) +
  ggplot2::geom_point() +
  ggplot2::facet_wrap(~ sector, scales = "free_y") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))

common_data %>%
  ggplot2::ggplot(ggplot2::aes(x = ef_diff, y = emission_factor.x)) +
  ggplot2::geom_point() +
  ggplot2::facet_wrap(~ sector, scales = "free_y") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))


```

In total, `r round(share_rows_replicated * 100, 3)` % of the rows in the banks data set have been successfully replicated.

We now compare the sectoral production totals and weighted average emission_factor from the Banks EO and the PAMS-based replica.

```{r compare_total}

min_year <- min(comparison_banks_eo$year)
max_year <- max(comparison_banks_eo$year)

totals_banks_eo <- comparison_banks_eo %>% 
  dplyr::summarise(
    emission_factor = stats::weighted.mean(.data$emission_factor, w = .data$production, na.rm = TRUE),
    production = sum(.data$production, na.rm = TRUE),
    .by = c("year", "sector")
  ) %>% 
  dplyr::filter(.data$year %in% c(min_year, max_year))

totals_pams_replica <- pams_to_banks_eo_complete %>% 
  dplyr::summarise(
    emission_factor = stats::weighted.mean(.data$emission_factor, w = .data$production, na.rm = TRUE),
    production = sum(.data$production, na.rm = TRUE),
    .by = c("year", "sector")
  ) %>% 
  dplyr::filter(.data$year %in% c(min_year, max_year))

sector_totals <- totals_banks_eo %>% 
  dplyr::inner_join(
    totals_pams_replica,
    by = c("sector", "year"),
    suffix = c("_banks", "_replica")
  ) %>% 
  dplyr::mutate(
    emission_factor_banks = dplyr::if_else(is.nan(.data$emission_factor_banks), NA_real_, .data$emission_factor_banks),
    emission_factor_replica = dplyr::if_else(is.nan(.data$emission_factor_replica), NA_real_, .data$emission_factor_replica)
  ) %>% 
  dplyr::mutate(
    prod_equal = dplyr::if_else(.data$production_banks == .data$production_replica, TRUE, FALSE),
    ef_equal = dplyr::if_else(.data$emission_factor_banks == .data$emission_factor_replica, TRUE, FALSE)
  ) %>% 
  dplyr::relocate("sector", "year", "production_banks", "production_replica", "prod_equal", "emission_factor_banks", "emission_factor_replica", "ef_equal")

DT::datatable(sector_totals)

```



## Conclusion

If all indicators in the table above are `TRUE`, the data set passes the consistency check. 

If some are not `TRUE`, please check:

- are there `NA` values, because of expected missing values? (e.g. `emission_factor` in technology pathway sectors)
- are the of the `FALSE` indicators likely due to rounding? (i.e. the actual summary values seem almost identical and the plots above indicate extremely low differences)

If the answer to either of the above points is "no", further investigation is needed.
