---
title: "Status Report - PAMS Comparison"
author: "Jackson Hoffart"
date: \today
output: 
  html_document:
      toc: true
      toc_depth: 2
      theme: flatly
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Error', '\u2716 **Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Warning:', '\u2716 **Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('\n\n<div class="alert alert-success">',
           gsub('##', '\n \u2714', x),
           '</div>', sep = '\n')
   }
)


config <- config::get()

# set configs
base_pams_path <- config$paths$path_pams_base
comparison_pams_path <- config$paths$path_pams_compare
company_id_threshold <- config$thresholds$company_id_threshold
sectoral_production_threshold <- config$thresholds$sectoral_production_threshold

library(dplyr)
library(tibble)
library(readxl)
library(tidyr)
library(glue)
library(pacta.data.preparation)
```

## Introduction

This report compares two input PAMS datasets: `base_pams` and `comparison_pams` to ensure quality assurance. We will test various expectations between the datasets to identify any unexpected differences or inconsistencies.

Some differences are of course expected between deliveries, which is why we introduce threshold limits. Thresholds can be set up-front in the `config.yml` file. 

## Load Data

The PAMS dataset has changed format between 2022Q4 and 2021Q4. Because of this, the function `pacta.data.preparation::import_ar_advanced_company_indicators()` will only work on the latest 2022Q4 data. The main missing sheet between these releases is the "Company Ownership" sheet. As such, we will create a slight variation of this import function that doesn't compare the missing sheet.

When using this workflow in the future, it should be explored if comparison metrics should be included to assess differences in the "Company Ownership" sheet. 

```{r import, echo = FALSE}
import_ar_advanced_company_indicators_no_ownership <- function(filepath, 
                                                               drop_nas = TRUE, 
                                                               fix_names = FALSE) {
  
  sheet_names <- readxl::excel_sheets(path = filepath)
  
  stopifnot(
    "Company Information" %in% sheet_names,
    # "Company Ownership" %in% sheet_names,
    "Company ISINs" %in% sheet_names,
    "Company Emissions" %in% sheet_names,
    "Company Activities" %in% sheet_names
  )
  
  company_info <-
    readxl::read_excel(path = filepath, sheet = "Company Information")
  
  # company_ownership <-
  #   readxl::read_excel(path = filepath, sheet = "Company Ownership") %>%
  #   filter(.data$`Ownership Level` == 0) %>%
  #   select(.data$`Company ID`, .data$`Is Ultimate Listed Parent`)
  
  company_isins <-
    readxl::read_excel(path = filepath, sheet = "Company ISINs") %>%
    group_by(.data$`Company ID`) %>%
    summarise(ISINs = list(.data$ISIN))
  
  company_emissions <-
    readxl::read_excel(path = filepath, sheet = "Company Emissions") %>%
    tidyr::pivot_longer(
      cols = dplyr::matches(" [0-9]{4}$"),
      names_to = c("consolidation_method", "year"),
      names_pattern = "^(.*) (20[0-9]{2})$",
      names_transform = list("year" = as.integer),
      values_to = "value",
      values_ptypes = numeric(),
      values_drop_na = drop_nas
    ) %>%
    mutate(value_type = "emission_intensity", .before = "year")
  
  company_activities <-
    readxl::read_excel(path = filepath, sheet = "Company Activities") %>%
    pivot_longer(
      cols = dplyr::matches(" [0-9]{4}$"),
      names_to = c("consolidation_method", "year"),
      names_pattern = "^(.*) (20[0-9]{2})$",
      names_transform = list("year" = as.integer),
      values_to = "value",
      values_ptypes = numeric(),
      values_drop_na = drop_nas
    ) %>%
    mutate(value_type = "production", .before = "year")
  
  output <-
    company_info %>%
    # dplyr::full_join(company_ownership, by = "Company ID") %>%
    dplyr::full_join(company_isins, by = "Company ID") %>%
    dplyr::right_join(
      dplyr::bind_rows(company_emissions, company_activities),
      by = c("Company ID", "Company Name")
    ) %>%
    mutate(across(where(is.character), as.factor))
  
  if (fix_names) {
    names(output) <- gsub(" ", "_", tolower(names(output)))
  }
  
  output
}

base_pams <- import_ar_advanced_company_indicators_no_ownership(base_pams_path)
comparison_pams <- import_ar_advanced_company_indicators_no_ownership(comparison_pams_path)
```

## Data Overview

Provide a brief overview of both datasets:

``` {r summary, echo = FALSE}

# Overview of base_pams
cat("Summary of base_pams:\n")
print(head(base_pams))

# Overview of comparison_pams
cat("\nSummary of comparison_pams:\n")
print(head(comparison_pams))
```

## Data Comparison
We will consider expectations across a variety of axes.

### Expectation 1: Company IDs
We expect not to lose or gain more than `r company_id_threshold`% of the total original number of Company IDs.

``` {r company_ids, echo = FALSE}
# Company IDs

# compute unique Company IDs
base_ids <- unique(base_pams$`Company ID`)
compare_ids <- unique(comparison_pams$`Company ID`)

# compute the number of IDs lost and gained
base_count <- length(base_ids)
lost_count <- length(setdiff(base_ids, compare_ids))
lost_percent <- 100 *(1 - ((base_count-lost_count)/base_count))

gained_count <- length(setdiff(compare_ids, base_ids))
gained_percent <- 100 *(1 - ((base_count-gained_count)/base_count))

# state statistics
if (gained_percent > company_id_threshold) {
  warning(glue::glue("PAMS gained {round(gained_percent, 2)}% new Company IDs"))
} else {
  message(glue::glue("PAMS gained {round(gained_percent, 2)}% new Company IDs"))
}

if (lost_percent > company_id_threshold) {
  warning(glue::glue("PAMS lost {round(lost_percent, 2)}% of original Company IDs"))
} else {
  message(glue::glue("PAMS lost {round(lost_percent, 2)}% of original Company IDs"))
}

```

### Expectation 2: Sectoral Production Values

We expect that production values, in aggregate, shouldn't differ by more than `r sectoral_production_threshold`%.

``` {r sectoral_production, echo = FALSE}
# Calculate the total production values for both datasets
total_production_base <- base_pams %>% 
  filter(`Is Ultimate Parent` == TRUE) %>% 
  filter(consolidation_method == "Direct Ownership") %>% 
  filter(value_type == "production") %>% 
  group_by(`Asset Sector`) %>% 
  summarize(value = sum(value))

total_production_comparison <- comparison_pams %>% 
  filter(`Is Ultimate Parent` == TRUE) %>% 
  filter(consolidation_method == "Direct Ownership") %>% 
  filter(value_type == "production") %>% 
  group_by(`Asset Sector`) %>% 
  summarize(value = sum(value))

# Check if the expectation is met
sector_expectation <- full_join(
  total_production_base, 
  total_production_comparison, 
  by = "Asset Sector",
  suffix = c("_base", "_comparison")
) %>% 
  mutate(
    difference = 100 * abs(value_base - value_comparison) / value_base
  )

# Summary
cat("\nExpectation 2: Production Values\n")
for (sector in unique(sector_expectation$`Asset Sector`)) {
  sector_difference <- filter(sector_expectation, `Asset Sector` == sector) %>% pull(difference)
  
  if (sector_difference > sectoral_production_threshold) {
    warning(glue::glue("PAMS total {sector} Production changed by {round(sector_difference, 2)}%"))
  } else {
    message(glue::glue("PAMS total {sector} Production changed by {round(sector_difference, 2)}%"))
  }
}

```

### Expectation 3: Sectors

We expect that the total possible options for Sector and Technology are identical between the two datasets. If there are any gained or dropped levels here, we want to know about it. 

``` {r sectors}
sectors_base <- unique(base_pams$`Asset Sector`)
sectors_compare <- unique(comparison_pams$`Asset Sector`)

lost_sectors <- setdiff(sectors_base, sectors_compare)
gained_sectors <- setdiff(sectors_compare, sectors_base)

if (length(lost_sectors) != 0) {
  warning(glue::glue("We lost the following sectors: {lost_sectors}"))
} 

if (length(gained_sectors) != 0) {
  warning(glue::glue("We gained the following sectors: {gained_sectors}"))
}

if (length(gained_sectors) == 0 && length(lost_sectors)) {
  message("Sectors are identical between datasets")
}

```

### Expectation 4: Technologies

``` {r technologies, echo = FALSE}
# this expectation assumes no lost or gained sectors
sectors_base <- unique(base_pams$`Asset Sector`)

for (sector in sectors_base) {
  technologies_base <- base_pams %>% 
    filter(`Asset Sector` == sector) %>% 
    distinct(`Asset Technology`) %>% 
    pull(`Asset Technology`)
  
  technologies_compare <- comparison_pams %>% 
    filter(`Asset Sector` == sector) %>% 
    distinct(`Asset Technology`) %>% 
    pull(`Asset Technology`)
  
  lost_technologies <- setdiff(technologies_base, technologies_compare)
  gained_technologies <- setdiff(technologies_compare, technologies_base)
  
  if (length(lost_technologies) != 0) {
    warning(glue::glue("In the {sector} Sector, we lost the following technologies: {glue_collapse(lost_technologies, ', ', last = ' and ')}"))
  }
  
  if (length(gained_technologies) != 0) {
    warning(glue::glue("In the {sector} Sector, we gained the following technologies: {glue_collapse(gained_technologies, ', ', last = ' and ')}"))
  }
  
  if (length(gained_technologies) == 0 && length(lost_technologies) == 0) {
    message(glue::glue("In the {sector} Sector, technologies are identical."))
  }
}

```

## Conclusion
In conclusion, this status report has compared the two datasets, `base_pams` and `comparison_pams`, based on the specified expectations. The summary of each expectation indicates whether it was met or not. If expectations were not met, details of the differences are provided.
